Supplementary Materials: Separating random and deterministic sources of computational noises in explore-exploit decisions

### 2024-05-14 manuscript/2024 - Plot Comp Bio/manuscript_DeterministicRandomNoise.tex
u

\subsubsection*{Model-based results}

Posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ are shown in Figure \ref{fig:mb1} and Supplementary Figure S8. Consistent with our model-free results, we see that both random and deterministic noise are non-zero. \replaced[id=siyu]{Numerically,}{and that} random noise is about 2-3 times larger than the deterministic noise. \added[id=siyu]{By computing the posterior distribution of $\sigma^2_{det}/(\sigma^2_{det}+\sigma^2_{ran})$, our data suggests that 14.25\% of the variability in random exploration is accounted for by deterministic noise ([4.90\%, 28.81\%], 95\% CI).} In addition, we find that both random and deterministic noise increase with horizon. This increase was larger for random noise (mean = 7.13, 100\% of samples showed an increase in random noise with horizon) than deterministic noise (mean = 2.59, 98.64\% of samples showed an increase in deterministic noise with horizon). But intriguingly, the relative increase in both types of noise was \replaced[id=siyu]{similar}{the same} (Figure \ref{fig:ratio}). That is, when we compute the relative increase in deterministic noise with horizon, $\sigma^{det}_{horizon6}/\sigma^{det}_{horizon1}$, it is \replaced[id=siyu]{very similar}{almost identical} to the relative increase in random noise with horizon $\sigma^{ran}_{horizon6}/\sigma^{ran}_{horizon1}$. 
%BOB SAYS: CHECK NOTATION HERE AND INSERT FIGURE SHOWING RATIO

\begin{figure}[H]
\begin{center}
	\includegraphics[width=0.7\textwidth]{figures/RDBayes_hyperprior.jpg}
	\caption{\added[id=bob]{TITLE ON B AND D SHOULD SAY `CHANGE IN RANDOM NOISE'}Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise. Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).  However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
	%BOB SAYS: DON'T NEED THE NUMBERS ON THE Y-AXES. 
	\label{fig:mb1}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
	\includegraphics[width=1\textwidth]{figures/RDBayes_2noise_modelA.jpg}
	\caption{
		% BOB SAYS: KEEP THE TOP LINE OF THIS FIGURE FOR THE MAIN PAPER AND MOVE THE FULL FIGURE TO SUPPLEMENTARY MATERIAL TO FIT DESCRIPTION IN TEXT.  
		% ALSO MAKE THE FOLLOWING CHANGES: THIS FIGURE NEEDS A LEGEND IN TOP RIGHT TO SAY WHICH LINE IS MODEL AND WHICH IS NOT.  I WOULD ALSO USE TEXTBOXES (in Matlab annotation('textbox', [0 0 1 1], 'string', 'hi') will get you started) TO DESCRIBE THE MODELS RATHER THAN LABELLING MODEL A ETC ... COULD EVEN HAVE SOME FORM OF TABLE 1 ON THE LEFT HAND SIDE - SO A COLUMN FOR DETERMINISTIC NOISE AND A COLUMN FOR RANDOM NOISE.
		Our model accounts for all qualitative patterns of the data, namely, p(high info) and p(low mean) increase as a function of horizon, p(inconsistent) increases as a function of horizon for both [1 3] and [2 2] conditions and lies between the pure random and pure deterministic noise prediction.}
	\label{fig:mb3}
\end{center}
\end{figure}

\subsubsection*{Comparison with alternative models}
COULD POTENTIALLY MOVE THIS TO THE SUPPLEMENT

To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model. These models varied \added[id=siyu]{in} whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon. \added[id=siyu]{Specifically, we tested the following 6 models (Note that the $\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$ model is our original full model).}

\begin{table}[h]
\small
\centering
\begin{tabular}{|c|c|c|}
	\hline
	Model & Deterministic noise & Random noise \\
	\hline
	$\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$  & Horizon dependent & Horizon dependent\\
	\hline
	$\sigma^{ran}_{horizon},\sigma^{det}_{}$ & Fixed & Horizon dependent\\
	\hline
	$\sigma^{ran}_{},\sigma^{det}_{horizon}$ & Horizon dependent & Fixed\\
	\hline
	$\sigma^{ran}_{},\sigma^{det}_{}$ & Fixed & Fixed\\
	\hline
	$\sigma^{ran}_{horizon}$ & Horizon dependent & None\\
	\hline
	$\sigma^{det}_{horizon}$ & None & Horizon dependent\\
	\hline
\end{tabular}
\caption{Variants of the model.}
\label{tab:models}	
\end{table}
\added[id=siyu]{
The posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ (when they exist) in these model variants are shown in Supplementary Figure S9. We again simulated choices using fitted parameters from these models and repeated the model-free analysis on the simulated data. 
}As shown in Supplementary Figure S10, only one of these models, where random noise is horizon dependent but deterministic noise is not, can capture the full qualitative pattern of \replaced[id=siyu]{behavior}{responding}. However, the quantitative fit to the data is not as good (Supplementary Figure S10).% SIYU says: they actually look quite similar, without horizon dependent deterministic noise, the fixed det noise model is only marginally worse on p(inconsistent) in [2 2] condition, as it overlaps with the pure random noise line more.

\added[id=siyu]{
Moreover, we examined if our model can indeed qualitatively capture whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon. To test this, we simulated choices from each of the 6 models, and then fit the simulated choices with our original full model. The simulation was repeated 50 times for each model. Indeed, we showed that our model can capture both the existence of random and deterministic noise, and whether each noise changes with horizon condition (Figure \ref{fig:6model}), with only one exception that our model falsely detected a small fraction of deterministic noise when no deterministic noise was present (Figure \ref{fig:6model}, S). This phenomenon was also examined and discussed in the section "Model validation" above.
}

\begin{figure}[H]
\begin{center}
	\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_hyperprior_6model_recovery.jpg}
	\caption{
		Our model qualitatively captures whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon. A-D. both deterministic and random noise are horizon dependent, E-H. only random noise is horizon dependent, I-L. only deterministic noise is horizon dependent, M-P. neither random nor deterministic noise is horizon dependent, Q-T. only deterministic noise is assumed to be present, U-X. only random noise is assumed to be present.}
	\label{fig:6model}
\end{center}
\end{figure}

%The previous section suggests that behavioral variability in random exploration can be at least partially explained by deterministic processing of the stimulus.

@@

\section*{Discussion}

% BOB SAYS: CAN YOU TRY TO REFRAME THIS AROUND THE DETERMINISTIC STORY
% \added[id=siyu]{Randomness in exploratory behavior can be driven by both the intrinsic variability generated in the brain (appearing in behavior as random noise) and by deterministic neural processes induced by the stimuli (appearing in behavior as deterministic noise).} 
In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise.'  Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for around 15\% of the overall variability in people's behavior.

%	\added[id=siyu]{We quantified the contribution of both deterministic noise (unobserved deterministic processes of the stimulus) and random noise (intrinsic stochasticity in decision making, or stimulus irrelevant processes) contribute to behavioral variability in random exploration,

SUGGEST REFRAMING NEXT PARAGRAPH IN THE FOLLOWING WAY.  SIMPLE READ OF RESULTS SUGGESTS RANDOM EXPLORATION IS RANDOM - THIS IS CONSISTENT WITH X.  FOLLOWING PARA THEN GIVES LIMITATIONS AND TALKS ABOUT DETERMINISTIC INTERPRETATION.  I'VE TRIED TO DO THIS ...

\added[id=bob]{One interpretation for this low level of deterministic noise is that most of the variability in random exploration is truly random. Such a random noise interpretation, would be consistent with recent work showing that variability in perceptual decisions may be driven by imperfections in mental inference \cite{drugowitsch16}. In this view, apparently random behavior is not due to sensory processing or response selection, but to suboptimal computations in the brain. Although suboptimal inference is different from simply adding random noise to neural circuitry\citep{Pouget12}, as long as the suboptimality in neural computation is not a deterministic function of the stimuli, it is a form of random noise in our definition. Indeed, a strong interpretation of this hypothesis would suggest that randomness in explore-exploit behavior is due to imperfect inference about the correct course of action. In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute CITE DEEP EXPLORATION PAPER.}

While the random noise interpretation is theoretically appealing, our approach, while an improvement on previous methods, is not without limitations. Most important is that our measure of `random' noise is only an upper bound on the true level of randomness and that, in principle, the random decision noise could be lower. Specifically, in our model, what we labeled random noise was really `non-stimulus-driven variability.' While this non-stimulus-driven variability could be driven by truly random stochastic processes, it could also be driven by deterministic processing that is unrelated to the stimuli in the task. For example, such deterministic noise could be driven by differences in where people look, or for how long they look, or by whether they were fidgeting or scratching their nose \citep{Musall2019}. In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:coverage2}, Supplementary Figure S6). As a result, from both a conceptual and methodological perspective, it is possible that the remaining 85\% of the decision noise that is not stimulus-driven noise, could be deterministic.

@@

To see how a change in reward processing could affect random and deterministic noise, consider the simple decision model we introduced in Equation XXX. In this model, choice is determined by the sign of the difference in utility $\Delta Q$ between the two options, where
\begin{equation}
	\label{eq:noise1}
	\Delta Q = 
	\Delta R + A \Delta I + b + n_{det} + n_{ran}
\end{equation}
Now imagine a case where the reward signal is scaled by a factor $\beta$.  In this case, $\Delta Q$ becomes
\begin{equation}
	\label{eq:noise2}
	\Delta Q = 
	\beta \Delta R + A \Delta I + b + n_{det} + n_{ran}
\end{equation}
Because the choice only depends on the sign of $\Delta Q$, scaling $\Delta Q$ by a factor of $1/\beta$ will not change the behavior of the model.  Thus, if we divide both sides of the above equation by $\beta$ we get
\begin{equation}
	\label{eq:noise3}
	\Delta Q / \beta = 
	\Delta R + A \Delta I / \beta + b / \beta + n_{det} / \beta + n_{ran} / \beta
\end{equation}
which is equivalent to a scaling of both deterministic and random noise by the same factor $1/\beta$. Thus, one interpretation of our result that both deterministic and random noise change across horizons with the same ratio, is that this reflects a change in reward processing. That is, the reward signal is reduced in the longer horizon condition (smaller $\beta$ in horizon 6 than horizon 1).  %In the alternate interpretation, where deterministic and random noise are varied independently it is not clear why they should change by the same scale factor.

Such a reduction in the strength of reward coding in exploration, is consistent with our recent work using a drift diffusion model (DDM) to model explore-exploit decisions CITE FENG ET AL. In the drift diffusion model, changes in behavioral variability can be driven by changes in the decision threshold (smaller threshold = more noise) or changes in the signal-to-noise ratio with which reward is encoded (lower SNR = more noise). By fitting both choices and response times, we were able to distinguish between these two accounts showing that the majority of the horizon-change in variability was driven by changes in SNR not threshold.  However, this model could not determine whether the changes in SNR were driven by signal or noise. By showing that the change in deterministic and random noise have the same ratio, the present work suggests that this SNR change is driven by changes in reward-signal processing, not noise. Of course, to truly see whether changes in signal or noise are driving random exploration will require more direct measurements of neural processing such as with neuroimaging and electrophysiology (CITE TOMOV GERSHMAN FMRI, CITE EBITZ).

@@

80 participants (ages 18-25, 37 male, 43 female) from the University of Arizona undergraduate subject pool participated in the experiment. 14 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}. In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved \deleted[id=siyu]{driven}by making a completely random choice between the two options and excluded participants with a likelihood smaller than 99.999\%, in other words, participants who didn't show an accuracy significant above chance with $p < 0.001$ were excluded in the analysis. This left 65 for the main analysis. Note that including the 15 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS

@@

\subsection*{Model-based analysis}
We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate \replaced[id=siyu]{deterministic noise from random noise.}{between components of the noise that are deterministically driven by the stimulus (`deterministic noise') and components of the noise that are not deterministically driven by the stimulus (`random noise').}Because the stimuli are identical in the repeated games, by definition, deterministic noise remains the same in repeated games, whereas random noise can change.

>>>>

\subsubsection*{Model-based results} 
Posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ are shown in Figure \ref{fig:mb1} and Supplementary Figure S11. Consistent with our model-free results, we see that both random and deterministic noise are non-zero. Numerically, random noise is about 2-3 times larger than the deterministic noise. By computing the posterior distribution of $\sigma^2_{det}/(\sigma^2_{det}+\sigma^2_{ran})$, our data suggests that 14.25\% of the variability in random exploration is accounted for by deterministic noise ([4.90\%, 28.81\%], 95\% CI). In addition, we find that both random and deterministic noise increase with horizon. This increase was larger for random noise (mean = 7.13, 100\% of samples showed an increase in random noise with horizon) than deterministic noise (mean = 2.59, 98.64\% of samples showed an increase in deterministic noise with horizon). But intriguingly, the relative increase in both types of noise was similar (Figure \ref{fig:ratio}). That is, when we compute the relative increase in deterministic noise with horizon, $\sigma^{det}_{horizon6}/\sigma^{det}_{horizon1}$, it is very similar to the relative increase in random noise with horizon $\sigma^{ran}_{horizon6}/\sigma^{ran}_{horizon1}$.

\begin{figure}[H]
\begin{center}
	\includegraphics[width=0.7\textwidth]{figures/RDBayes_hyperprior.jpg}
	\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise. Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).  However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
	%BOB SAYS: DON'T NEED THE NUMBERS ON THE Y-AXES. 
	\label{fig:mb1}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
	\includegraphics[width=1\textwidth]{figures/RDBayes_2noise_modelA.jpg}
	\caption{
		Our model accounts for all qualitative patterns of the data, namely, p(high info) and p(low mean) increase as a function of horizon, p(inconsistent) increases as a function of horizon for both [1 3] and [2 2] conditions and lies between the pure random and pure deterministic noise prediction.}
	\label{fig:mb3}
\end{center}
\end{figure}

@@

\section*{Discussion}
In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise.'  Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for around 15\% of the overall variability in people's behavior.

% SUGGEST REFRAMING NEXT PARAGRAPH IN THE FOLLOWING WAY.  SIMPLE READ OF RESULTS SUGGESTS RANDOM EXPLORATION IS RANDOM - THIS IS CONSISTENT WITH X.  FOLLOWING PARA THEN GIVES LIMITATIONS AND TALKS ABOUT DETERMINISTIC INTERPRETATION.  I'VE TRIED TO DO THIS ...

One interpretation for this low level of deterministic noise is that most of the variability in random exploration is truly random. Such a random noise interpretation, would be consistent with recent work showing that variability in perceptual decisions may be driven by imperfections in mental inference \cite{drugowitsch16}. In this view, apparently random behavior is not due to sensory processing or response selection, but to suboptimal computations in the brain. Although suboptimal inference is different from simply adding random noise to neural circuitry\citep{Pouget12}, as long as the suboptimality in neural computation is not a deterministic function of the stimuli, it is a form of random noise in our definition. Indeed, a strong interpretation of this hypothesis would suggest that randomness in explore-exploit behavior is due to imperfect inference about the correct course of action. In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute \citep{Wilson2020}.

Although the random noise interpretation is theoretically appealing, our approach, while an improvement on previous methods, is not without limitations. Most important is that our measure of `random' noise is only an upper bound on the true level of randomness and that, in principle, the random decision noise could be lower. Specifically, in our model, what we labeled random noise was really `non-stimulus-driven variability'. While this non-stimulus-driven variability could be driven by truly random stochastic processes, it could also be driven by deterministic processing that is unrelated to the stimuli in the task. For example, such deterministic noise could be driven by differences in where people look, or for how long they look, or by whether they were fidgeting or scratching their nose \citep{Musall2019}. In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:paramrecover_main}, Supplementary Figure S5, S6). As a result, from both a conceptual and methodological perspective, it is possible that the remaining 85\% of the decision noise that is not stimulus-driven noise, could be deterministic.

@@

To see how a change in reward processing could affect random and deterministic noise, consider the simple decision model we introduced in Equation \ref{eq:newmodel}. In this model, choice is determined by the sign of the difference in utility $\Delta Q$ between the two options, where
\begin{equation}
	\label{eq:noise1}
	\Delta Q = 
	\Delta R + A \Delta I + b + n_{det} + n_{ran}
\end{equation}
Now imagine a case where the reward signal is scaled by a factor $\beta$.  In this case, $\Delta Q$ becomes
\begin{equation}
	\label{eq:noise2}
	\Delta Q = 
	\beta \Delta R + A \Delta I + b + n_{det} + n_{ran}
\end{equation}
Because the choice only depends on the sign of $\Delta Q$, scaling $\Delta Q$ by a factor of $1/\beta$ will not change the behavior of the model.  Thus, if we divide both sides of the above equation by $\beta$ we get
\begin{equation}
	\label{eq:noise3}
	\Delta Q / \beta = 
	\Delta R + A \Delta I / \beta + b / \beta + n_{det} / \beta + n_{ran} / \beta
\end{equation}
which is equivalent to a scaling of both deterministic and random noise by the same factor $1/\beta$. Thus, one interpretation of our result that both deterministic and random noise change across horizons with the same ratio, is that this reflects a change in reward processing. That is, the reward signal is reduced in the longer horizon condition (smaller $\beta$ in horizon 6 than horizon 1).  %In the alternate interpretation, where deterministic and random noise are varied independently it is not clear why they should change by the same scale factor.

Such a reduction in the strength of reward coding in exploration, is consistent with our recent work using a drift diffusion model (DDM) to model explore-exploit decisions \citep{Feng21}. In the drift diffusion model, changes in behavioral variability can be driven by changes in the decision threshold (smaller threshold = more noise) or changes in the signal-to-noise ratio with which reward is encoded (lower SNR = more noise). By fitting both choices and response times, we were able to distinguish between these two accounts showing that the majority of the horizon-change in variability was driven by changes in SNR not threshold. However, this model could not determine whether the changes in SNR were driven by signal or noise. By showing that the change in deterministic and random noise have the same ratio, the present work suggests that this SNR change is driven by changes in reward-signal processing, not noise. Of course, to truly see whether changes in signal or noise are driving random exploration will require more direct measurements of neural processing such as with neuroimaging and electrophysiology \citep{Tomov20, Ebitz18,costa1,costa22}

@@

80 participants (ages 18-25, 37 male, 43 female) from the University of Arizona undergraduate subject pool participated in the experiment. 14 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}. In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved by making a completely random choice between the two options and excluded participants with a likelihood smaller than 99.999\%, in other words, participants who didn't show an accuracy significant above chance with $p < 0.001$ were excluded in the analysis. This left 65 for the main analysis. Note that including the 15 badly performing subjects did not change the main results (Supplementary Figures S1, S2, S11)

@@

\subsection*{Model-based analysis}
We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate deterministic noise from random noise. Because the stimuli are identical in the repeated games, by definition, deterministic noise remains the same in repeated games, whereas random noise can change.

### 2024-05-14 manuscript/2024 - Plot Comp Bio/supplemental/DetRanNoise_Supplementary.tex
u

	\section{Additional model-based analyses}
	\subsection{Replication of Figure 4 without excluding subjects}
	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=0.7\textwidth]{figures/RDBayes_hyperprior__all.jpg}
			\caption{Model based analysis with data from all participants (i.e. no exclusions) showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise. Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).  However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
			\label{fig:s8}
		\end{center}
	\end{figure}
	\newpage
	\subsection{Fitting variations of the two-noise model to the data}
	\begin{figure}[hp]
		\begin{center}
			\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_hyperpriors_6model.jpg}
			\caption{Model based analysis with variations of models. Each row is one model.  These models varied in whether deterministic $\sigma^{det}$ and random noise $\sigma^{ran}$ are present or not and whether either types of noise is dependent on horizon (subscript denotes the dependence on horizon). }
			\label{fig:s9}
		\end{center}
	\end{figure}

>>>>

	\section{Additional model-based analyses}
	\subsection{Replication of Figure 4 without excluding subjects}
	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=0.7\textwidth]{figures/RDBayes_hyperprior__all.jpg}
			\caption{Model based analysis with data from all participants (i.e. no exclusions) showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise. Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).  However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
			\label{fig:s8}
		\end{center}
	\end{figure}
	\newpage
	\subsection{Fitting variations of the two-noise model to the data}

	\subsubsection*{Comparison with alternative models}
	COULD POTENTIALLY MOVE THIS TO THE SUPPLEMENT

	To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model. These models varied \added[id=siyu]{in} whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon. \added[id=siyu]{Specifically, we tested the following 6 models (Note that the $\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$ model is our original full model).}

	\begin{table}[h]
		\small
		\centering
		\begin{tabular}{|c|c|c|}
			\hline
			Model & Deterministic noise & Random noise \\
			\hline
			$\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$  & Horizon dependent & Horizon dependent\\
			\hline
			$\sigma^{ran}_{horizon},\sigma^{det}_{}$ & Fixed & Horizon dependent\\
			\hline
			$\sigma^{ran}_{},\sigma^{det}_{horizon}$ & Horizon dependent & Fixed\\
			\hline
			$\sigma^{ran}_{},\sigma^{det}_{}$ & Fixed & Fixed\\
			\hline
			$\sigma^{ran}_{horizon}$ & Horizon dependent & None\\
			\hline
			$\sigma^{det}_{horizon}$ & None & Horizon dependent\\
			\hline
		\end{tabular}
		\caption{Variants of the model.}
		\label{tab:models}	
	\end{table}
	\added[id=siyu]{
		The posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ (when they exist) in these model variants are shown in Supplementary Figure S9. We again simulated choices using fitted parameters from these models and repeated the model-free analysis on the simulated data. 
	}As shown in Supplementary Figure S10, only one of these models, where random noise is horizon dependent but deterministic noise is not, can capture the full qualitative pattern of \replaced[id=siyu]{behavior}{responding}. However, the quantitative fit to the data is not as good (Supplementary Figure S10).% SIYU says: they actually look quite similar, without horizon dependent deterministic noise, the fixed det noise model is only marginally worse on p(inconsistent) in [2 2] condition, as it overlaps with the pure random noise line more.

	\added[id=siyu]{
		Moreover, we examined if our model can indeed qualitatively capture whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon. To test this, we simulated choices from each of the 6 models, and then fit the simulated choices with our original full model. The simulation was repeated 50 times for each model. Indeed, we showed that our model can capture both the existence of random and deterministic noise, and whether each noise changes with horizon condition (Figure \ref{fig:6model}), with only one exception that our model falsely detected a small fraction of deterministic noise when no deterministic noise was present (Figure \ref{fig:6model}, S). This phenomenon was also examined and discussed in the section "Model validation" above.
	}

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_hyperprior_6model_recovery.jpg}
			\caption{
				Our model qualitatively captures whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon. A-D. both deterministic and random noise are horizon dependent, E-H. only random noise is horizon dependent, I-L. only deterministic noise is horizon dependent, M-P. neither random nor deterministic noise is horizon dependent, Q-T. only deterministic noise is assumed to be present, U-X. only random noise is assumed to be present.}
			\label{fig:6model}
		\end{center}
	\end{figure}

	\begin{figure}[hp]
		\begin{center}
			\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_hyperpriors_6model.jpg}
			\caption{Model based analysis with variations of models. Each row is one model.  These models varied in whether deterministic $\sigma^{det}$ and random noise $\sigma^{ran}$ are present or not and whether either types of noise is dependent on horizon (subscript denotes the dependence on horizon). }
			\label{fig:s9}
		\end{center}
	\end{figure}
